Analytics Without Parameter Tuning Considered Harmful?

### 2015-03-11 dm101.tex
Updates from ShareLaTeX

Since programming inherently
introduces defects into  programs, is important to thoroughly {\em test} software before it is {\em used}.
Such testing  can be very expensive.
Software assessment budgets are finite
while assessment effectiveness increases 
exponentially with assessment effort.
Lowry et al. warn that  
the state space explosion problem imposes
strict limits on how much a system can be explored
via automatic formal methods~\cite{lowrey98}.
For standard black-box testing methods,
a {\em linear} increase
in the confidence $C$ that we have found all defects
can take {\em exponentially} more effort.
For example, for one-in-a-thousand detects,
moving $C$ from  
90\% to 98\% nearly doubles the number of required tests (2301 to   3910 black box
probes, respectively)\footnote{A randomly selected 
input to a program will find a fault with probability $p$.
After $N$ random black-box tests, the chances of the inputs 
not revealing any fault 
is $(1-p)^N$. Hence, the chances $C$ of seeing the fault is $1-(1-p)^N$
which can be rearranged to 
 $N(C,p)=\frac{log(1 -
C)}{log(1-p)}$. For example, $N(0.90,10^{-3})=2301$.}.

>>>>

Since programming inherently
introduces defects into  programs, is important to thoroughly {\em test} software before it is {\em used}.
Such testing  can be very expensive.
Software assessment budgets are finite
while assessment effectiveness increases 
exponentially with assessment effort.
Lowry et al. warn that  
the state space explosion problem imposes
strict limits on how much a system can be explored
via automatic formal methods~\cite{lowrey98}.
For standard black-box testing methods,
a {\em linear} increase
in the confidence $C$ that we have found all defects
can take {\em exponentially} more effort.
For example, for one-in-a-thousand detects,
moving $C$ from  
90\% to 98\% nearly doubles the number of  tests (2301 to   3910 black box
probes, respectively)\footnote{A randomly selected 
input to a program will find a fault with probability $p$.
After $N$ random black-box tests, the chances of the inputs 
not revealing any fault 
is $(1-p)^N$. Hence, the chances $C$ of seeing the fault is $1-(1-p)^N$
which can be rearranged to 
 $N(C,p)=\frac{log(1 -
C)}{log(1-p)}$. For example, $N(0.90,10^{-3})=2301$.}.

### 2015-03-11 tuningpredictor.tex
Updates from ShareLaTeX

\documentclass{sig-alternative}
\usepackage{multirow}
\usepackage{color}
\usepackage{graphics}
\usepackage{rotating}
\usepackage{graphics}
\usepackage{colortbl} 
\usepackage{picture}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\renewcommand{\footnotesize}{\scriptsize}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
%%% graph
\newcommand{\crule}[3][darkgray]{\textcolor{#1}{\rule{#2}{#3}}}
\newcommand{\rone}{\crule{1mm}{1.95mm}}
\newcommand{\rtwo}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}}
\newcommand{\rthree}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}}
\newcommand{\rfour}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}} 
\newcommand{\rfive}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}}
\newcommand{\quart}[3]{\begin{picture}(100,6)%1
{\color{black}\put(#3,3){\circle*{4}}\put(#1,3){\line(1,0){#2}}}\end{picture}}
\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

@@

\conferenceinfo{FSE}{'15 Bergamo, Italy}
\title{Implications of Differential Evolution for  Defect Prediction:
Analytics Without Parameter Tuning Considered Harmful?}
\numberofauthors{1}
\author{
\alignauthor
Wei Fu, Tim Menzies, Vivek Nair\\
       \affaddr{Computer Science, North Carolina State University, Raleigh, USA}\\
       {wfu@ncsu.edu, tim.menzies@gmail.com, vivekaxl@gmail.com}}

@@

\vspace{1mm}
\noindent
{\bf General Terms:} Experimentation, Algorithm

@@

@@

Many researchers have applied automatic optimizers to find good tunings.
This paper shows that for at least one software analytics problem
(software defect prediction from static code attributes),   a
very simple optimizer called differential evolution needs only   seconds
to just  few minutes to find good tunings.

These results   challenge much prior work into defect prediction:
Numerous   prominent papers have claimed 
that learnerX is better than learnerY for some software analytics task~\cite{lessmann2008benchmarking,hall11,me07b}.
For example, a recent IEEE TSE article claimed that the 
CART decision tree learner was far worse than Random Forests for
softwre defect predition~\cite{lessmann2008benchmarking}. 
Such conclusions do not survive tuning.
For example,
after tuning, the worst untuned learner (CART) can out-perform the supposedly
best learner (Random Forests).

Another class of conclusions challenged by this paper are research papers
that use data miners to   show that different attributes are ``best'' for different
data sets and even different goals for the tunings. Hence, it is no longer
supportable to use data mining to argue that certain kinds of attributes 
are better than others; e.g. as  done by~\cite{rahman2013how}. 
For example, in 2013,
Rahman et al
\cite{rahman2013how}  used results from logistic regression to conclude
that code metrics are generally less useful than process
metrics for  predicting defects from static code analysis.  
In the same year, Bell et al. (who also used negative binomial regression) argued that
software defect prediction was not improved by adding attributes relating to the social
structure of programmer teams~\cite{bell2013limited}.

Given the simplicity and effectiveness of differential evolution,
and also given the effects this method has on defect prediction, this
paper  motivates  a change to the entire field of software analytics.
At least for softwre defect prediction,
it is no longer enough to just run a data miner and present the result
{\em without} first conducting an tuning optimizations study.

The implications for other kinds of data mining tasks is unclear.
While this paper reports success with tuning defect predictors,
we were not so successful with
other analytics tasks (software development effort estimation).
Hence, we assert that it is now an open and pressing research question is whether or not  
tuning studies are now   required for research reports on software analytics.

>>>>

\documentclass{sig-alternative}
\usepackage{multirow}
\usepackage{color}
\usepackage{graphics}
\usepackage{cite}
\usepackage{rotating}
\usepackage{graphics}
\usepackage{colortbl} 
\usepackage{picture}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\renewcommand{\footnotesize}{\scriptsize}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
%%% graph
\newcommand{\crule}[3][darkgray]{\textcolor{#1}{\rule{#2}{#3}}}
\newcommand{\rone}{\crule{1mm}{1.95mm}}
\newcommand{\rtwo}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}}
\newcommand{\rthree}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}}
\newcommand{\rfour}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}} 
\newcommand{\rfive}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}}
\newcommand{\quart}[3]{\begin{picture}(100,6)%1
{\color{black}\put(#3,3){\circle*{4}}\put(#1,3){\line(1,0){#2}}}\end{picture}}
\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

@@

\conferenceinfo{FSE}{'15 Bergamo, Italy}
\title{ 
Analytics Without Parameter Tuning Considered Harmful?}
\numberofauthors{1}
\author{
\alignauthor
Wei Fu, Tim Menzies, Vivek Nair\\
       \affaddr{Computer Science, North Carolina State University, Raleigh, USA}\\
       fuwei.ee, tim.menzies, vivekaxl@gmail.com}

@@

@@

In the $21^{st}$ century, it is now impossible
to manually browse all the available software project
data. The PROMISE repository of SE data has grown to 200+ projects~\cite{promise15}
and this is just one of over a dozen open-source repositories
that are readily available to researchers~\cite{rod12}.
For example,  the time of this writing (Feb  2015), our web searches show that Mozilla Firefox has over 1.1 million bug reports, and platforms such as GitHub hosts over 14 million projects.

Faced with this information overload,
researchers in empirical SE
use  data miners  to (e.g.) generate 
defect predictors from static code measures.
Such   measures can be
automatically extracted from the code base, with very little effort
even for very large software systems~\cite{nagappan05}. 
In addition, they have  some generality
across multiple projects: e.g. defect
predictors developed at NASA~\cite{me07b} have also been successfully
applied in Turkey~\cite{tosun09}.
Such detectors reduce the effort required for 
defect prediction: if 
inspection teams let themselves be guided by defect predictors then
they can find 80\% to 88\% of the defects
after inspecting  20\% to 25\% of the code~\cite{ostrand04,tosun10}.

@@

Many researchers have applied automatic optimizers to find good tunings~\cite{cora10}  but not in the arean
of adjusting the parameters of data miners learning
static code defect predictors.
(to the best of our knowledge, this is the first such study).
This paper shows that for at least one software analytics problem
(software defect prediction from static code attributes),   a
very simple optimizer called 
{\em differential evolution} can  find good tunings.

The good news here is that that tuning defect predictors is very easy. Differential
evolution is so simple to implement that many
researchers just ``roll their own'' rather
than relying on some standard package.
By various measures
(presented below),  
tuning defect predictors is at least an order of magnitude
simpler than other standard optimization tasks. 
Further, measured in terms on runtimes, 
differental evolution takes just seconds
to a few minutes to propose highy effective tunings.

The not-so-good-news is that these results   challenge much prior work into defect prediction.
Several   prominent IEEE TSE papers~\cite{lessmann2008benchmarking,hall11,me07b} have claimed 
that learnerX is better than learnerY for some software analytics task~.
For example, a recent IEEE TSE article claimed that the 
CART decision tree learner was far worse than Random Forests for
softwre defect predition~\cite{lessmann2008benchmarking}. 
Such conclusions do not survive tuning.
For example,
after tuning, the worst untuned learner (CART) can out-perform the supposedly
best learner (Random Forests).

Another class of conclusions challenged by this paper are research papers
that use data miners to   show that certain factors
are more important than others for (say)
predicting defects. We show in this paper
that cuch conclusions
do not survive tuning. It turns out the after tuning,
different data sets show  that very different 
factors are ``best''. Also, even within the same data set, if we adjust the goals of the tuning then we arrive at different ``best'' factors.

Given this locality effect in what is ``best'',
then there are many   papers that may
need to be rewritten.
For example, one of us (Menzies) used data miners
to assert that some factors were more important than others for predicting
successful software reuse~\cite{me02k}. That assertion should now be doubted.

Similarly,   we should now revisit
all other conclusions that used data miners to rank different factors including:
\bi
\item
Bell et al.~\cite{bell2013limited} who say that
software defect prediction is not improved by adding attributes relating to the social
structure of programmer teams;
\item
Rahman et al
\cite{rahman2013how}  who say
 code metrics are  less useful than process
metrics for  predicting defects from static code analysis; 
\item Mosser et al.~\cite{moser2008comparative} who also say  that process metrics are better. 
\ei
There are many other examples of researchers
using data mining to claim that certain
factors are more important than others; e.g. ~\cite{zimmermann2007predicting,herzig2013predicting}.
All those results must  now be called into question.
At least for softwre defect prediction,
since is no longer enough to just run a data miner and present the result
{\em without} first conducting an tuning optimizations study.

The implications for other kinds of data mining tasks is unclear.
While this paper reports success with tuning defect predictors,
there are many of software analytics tasks 
(software development effort estimation, social network mining,
detecting duplicate issue reports, etc).
Since those other tasks often use the same kinds of learners
explored in this paper, it is clearly possible that
the conclusions of this paper apply to other tasks as well. 
Hence, we assert that it is now an open and pressing research question if
``analytics without parameter tuning is considered harmful?''.
Note that if the answer is ``yes'' then:
conferences like FSE
should routinely reject software analytics papers that do not
have a tuning study.

The rest of this paper is structured as follows.

One caveat before beginning- this paper explores
optimization via differential evolution.
XXX
